Text Is All You Need

### 2023-05-10 docs/hes_final/presentation.qmd
Transformer Slides First Pass

- Our third approach used a pair of encoder-only transformers, but for simplicity, we will refer to the first of these models as “the encoder” and the second as the “classifier”.
- The purpose of the encoder was to reduce the dimensionality of the embedding corpus by analyzing a subset of article embeddings and returning a single vector of substantially reduced size. This is roughly analogous to the role of a convolutional layer in a CNN. 
- The classifier could then be used to make final predictions on the resulting, distilled dataset.

@@

- To prepare the training dataset, our full sample of the GDELT corpus was split into a single $512\times n$ matrix for each country and month where $n$ is the number of articles published in that month related to that country.
- Each of these matrices was then split into several random subsamples of 50 articles each. 
- Samples for countries in which $n<50$ were padded to $512\times50$.

@@

![Reducing Embedding Dimensionality with Transformer Encoder](imgs/Transformer_Viz.png){#fig-transformer-viz}

- Once training was complete, the final layer of the encoder was removed, exposing the penultimate layer of 16 dense neurons. 
- The output of these 16 neurons would be used in the classification stage.

- With the encoder's training complete, we returned to the original base dataset. Each country-month matrix was again split into $512\times50$ matrices.
- Each of these samples could then be passed into the encoder, which distilled it into a $length-16$ vector.
- Each vector was then stacked, resulting in a single training example for each country and month of size $16\times\frac{n}{50}$.

>>>>

- Our third approach used a pair of encoder-only transformers
- the “encoder” and the “classifier”.
- The encoder was used to reduce the dimensionality of the embedding corpus by analyzing a subset of article embeddings and returning a single vector of substantially reduced size.
- Analogous to the role of a convolutional layer in a CNN. 
- The classifier could then be used to make final predictions on the resulting, distilled dataset.

@@

:::: {.columns}

::: {.column width="50%"}
![Transformer Model](imgs/Transformer_Viz.png){#fig-transformer-viz}
:::

::: {.column width="50%"}
Training the Encoder

- Each country/month sample was split into subsamples of 50 articles each
- The same label was used for each subsample, creating multiple unique training samples per country per month 
- This produced a larger, but more distorted corpus
:::

::::

@@

:::: {.columns}

::: {.column width="50%"}
![Transformer Model](imgs/Transformer_Viz.png){#fig-transformer-viz}
:::

::: {.column width="50%"}
Training the Classifier

- Once training was complete, the final layer of the encoder was removed, exposing the penultimate layer of 16 dense neurons. 
- Each vector was then stacked, resulting in a single training example for each country and month of size $16\times\frac{n}{50}$. 
:::

::::
